---
title: "Group_Assignment_2"
author: "Rahul Jadhav"
date: "2/7/2022"
output: 
  html_document:
     theme: cerulean
     toc: true
     toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages<-(c("dplyr","psych","ggplot2","fastDummies","broom","gtools", "scales","skimr","car","caret", "stargazer","effects"))
package.check <- lapply(
  packages,
  function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
lapply(c(packages),require, character.only = TRUE)


```

## Data Preparation

Read the Dataset using read.csv command.The main_dataset is the dataset which contains all the observation.
Checked for Na values in dataset and removed all the Na values.
Used random sampling and developed 5 sampled dataset with each having no similar observations.

```{r}
main_dataset <- read.csv('Chicago_Crimes_2001_to_2017_Part2.csv')

## Assuming homicide, Assault, robbery, and sexual offense should be the categories for violent crimes. Let's put these columns together in a group.
unique(main_dataset$Primary.Type)
cleaned_dataset <- main_dataset %>%
  mutate(
    Primary.Type = ifelse(Primary.Type=="HOMICIDE","HOMICIDE",
                          ifelse(Primary.Type=="CRIM SEXUAL ASSAULT" | 
                                   Primary.Type=="SEX OFFENSE" , "SEXUAL OFFENCE",
                                 ifelse(Primary.Type == "ASSAULT","ASSAULT",
                                        ifelse(Primary.Type == "ROBBERY", "ROBBERY","Other"
                                               )
                                        )
                                 )
                          )
  )

cleaned_dataset <- filter(cleaned_dataset,cleaned_dataset$Primary.Type=="SEXUAL OFFENCE" | 
                            cleaned_dataset$Primary.Type=="ASSAULT" | 
                            cleaned_dataset$Primary.Type=="HOMICIDE" | 
                            cleaned_dataset$Primary.Type=="ROBBERY" )




cleaned_dataset <- cleaned_dataset %>%
  select(Date,Block,Primary.Type,Location.Description,Arrest,Domestic,District,Ward,Year,Parts)

sapply(cleaned_dataset, function(x) sum(is.na(x)))


## convertion of few variables and data variable class from character to POSIXct (data-time)
## Since arrest column had unique values as True, TRUE, False and FALSE. Replaced True with TRUE and False with FALSE

cleaned_dataset <- cleaned_dataset %>%
  mutate(
    Arrest = ifelse(
      Arrest == "False", "FALSE", 
      ifelse(
        Arrest == "True", "TRUE", cleaned_dataset$Arrest
      )),
    Domestic = ifelse(
      Domestic == "False", "FALSE", 
      ifelse(
        Domestic == "True", "TRUE", cleaned_dataset$Domestic
      )),
    Date = as.POSIXlt(Date, format = "%m/%d/%Y %I:%M:%S %p")
    )

cleaned_dataset <- cleaned_dataset %>%
  mutate(
    Month = format(cleaned_dataset$Date, format="%m"),
    Day = format(cleaned_dataset$Date, format="%d"),
    Time = format(cleaned_dataset$Date, format="%H")) %>%
  select(-Date)

## converting all Qualitative data to factorial Data
cleaned_dataset[sapply(cleaned_dataset, is.character)] <- lapply(
  cleaned_dataset[sapply(
    cleaned_dataset, 
    is.character)]
  ,as.factor)



```
## SUMMARY STATISTIC

For numerical variable, we have created subset of numerical variable and categorical variable.


```{r}
## Summary Statistic of Continuous variables
psych::describe(cleaned_dataset) %>%
  dplyr::select(count = n,
         Mean = mean,
         Standard_Deviation = sd,
         )%>%
  mutate(
    Coefficient_Variation = Standard_Deviation/Mean
  )%>% na.omit()


## Summary Statistic of Quantiative Variables
summary(cleaned_dataset[sapply(cleaned_dataset , is.factor)])

```



## BASIC ANALYSIS
### Primary type of crimes 

```{r}

group_by_primary_type <- cleaned_dataset %>%
  group_by(Primary.Type,Domestic) %>%
  summarise(count = n())%>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange(desc(percentage))


ggplot(group_by_primary_type,aes(x=Primary.Type,y=count,fill=Domestic))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(label = comma)+
  ggtitle("Bar Plot : 01 Top 10 crimes reported in Chicago with Domestic Violence")

group_by_primary_type <- cleaned_dataset %>%
  group_by(Primary.Type,Arrest) %>%
  summarise(count = n())%>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange(desc(percentage)) 


ggplot(group_by_primary_type,aes(x=Primary.Type,y=count,fill= Arrest))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(label = comma)+
  ggtitle("Bar Plot : 02 Crimes reportd in chicago with Arrested Data")


```

### TimeBased Crime rate
```{r,figures-side, fig.show="hold", out.width="50%", results='hide', comment = " ", }

group_by_Year <- cleaned_dataset %>%
  group_by(Year,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange((Year))


ggplot(group_by_Year,aes(x=Year,y=count, fill= Domestic))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust=1))+
  ggtitle("Bar Plot : 03 Yearly Crime rate of Chicago from 2001 to 2017")


group_by_Month <- cleaned_dataset %>%
  group_by(Month,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange((Month))


ggplot(group_by_Month,aes(x=Month,y=count, fill= Domestic))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust=1))+
  ggtitle("Bar Plot : 03 Monthly Crime rate of Chicago from 2001 to 2017")

group_by_Day <- cleaned_dataset %>%
  group_by(Day,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange((Day))


ggplot(group_by_Day,aes(x=Day,y=count, fill= Domestic))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust=1))+
  ggtitle("Bar Plot : 03 Monthly Crime rate of Chicago from 2001 to 2017")

group_by_Time <- cleaned_dataset %>%
  group_by(Time,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange((Time))


ggplot(group_by_Time,aes(x=Time,y=count, fill= Domestic))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust=1))+
  ggtitle("Bar Plot : 03 Monthly Crime rate of Chicago from 2001 to 2017")


```

### District Crime rate
```{r}
group_by_District <- cleaned_dataset %>%
  group_by(District,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  )
ggplot(group_by_District,aes(x=District,y=count, fill= Domestic))+
  geom_bar(stat="identity")+
  ggtitle("Bar Plot : 05 District wise cases")+
  xlab("District")

```


## CHI-SQUARE ANALYSIS
Performing Chi-Square test to examine if Sale Price varies by Neighborhood

H0 (null hypothesis)        -> There is no relationship between District and Primary type of Crime
H1 (Alternative Hypothesis) -> There is relationship between District and Primary type of Crime
```{r}
chi <- cleaned_dataset
myChisq <- xtabs(~ as.factor(Primary.Type) + as.factor(District), data = chi)

tidy(chisq.test(myChisq))%>%
  select(
    method,
    DF = parameter,
    p.value,
    statistic
  )%>%
  mutate(
    statistic = round(statistic,2)
  )
```

Interpretation :
A chi-square test was performed to examine the relation between District and Primary type of crimes in Chicago. The relation between these variables was significant, X^2 (1,N=66) = 2704.5, p = 0. 

Since the p.value is very significant, we will reject the null hypothesis.


## 2 WAY ANOVA TEST 
Performing ANOVA test to examine if District varies by Primary type of crime and Arrests.

H0(null hypothesis) -> 
  * Primary type of crime and Arrest has no effect on District. Also there is no interaction affect on District due to Primary type of crime and Arrests.
  
H1(Alternative Hypothesis) ->
  * Primary type of crime and Arrest has effect on District. Also there is interaction affect on District due to Primary type of crime and Arrests.

```{r }
two_way_fit <- aov(District ~as.factor(Primary.Type) * as.factor(Arrest) , data=cleaned_dataset)
tidy(two_way_fit)%>%
  mutate(
    sumsq = signif(sumsq, digits = 2),
    p.value = round(p.value,3),
    statistic = round(statistic,3)
    
  )

```

Interpretation :
A two-way ANOVA was performed to analyze the effect of Primary type of crime and Arrest on District.
A two-way ANOVA revealed that -
Simple main effects analysis showed that Primary type of crime did have a statistically significant effect on District (p = 0).

 
Simple main effects analysis showed that Arrest did have a statistically significant effect on Sale Price (p = 0).

## INTERACTION PLOT FROM ANOVA


```{r, comment = " ", fig.width=22, fig.height=7, fig.align='center'}
## merging Exterior values lesser than 60 
test <- group_by_primary_type <- cleaned_dataset %>%
  group_by(Primary.Type,Domestic,District) %>%
  summarise(count = n())%>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange(desc(percentage))
interaction.plot(x.factor = test$District,
                 trace.factor = test$Primary.Type,
                 response = test$count,
                 fun = mean,
                 type = "b",
                 ylab = "Count",
                 xlab = "District",
                 col = c("red","green","blue","black"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "Primary type crimes",
                 pch=c(19),             ### Symbols for levels of trace var.
                 fixed=TRUE,                    ### Order by factor order in data
                 leg.bty = "o")
 
```


## Spliting Dataset into train and test sets

Here we are going to sample our main dataset into 70-30 ratio as Train and test dataset respectively.
```{r}
set.seed(123)
trainIndex <- createDataPartition(cleaned_dataset$Arrest, p= 0.70, list = FALSE)
train <- cleaned_dataset[trainIndex,]
test <- cleaned_dataset[-trainIndex,]
```

## Linear Regression Model 
Here we are going to check if crime rate has anything to do with the District, Arrest Made.

```{r}
Lm_train_dataset <- train %>%
  group_by(District, Arrest,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    Arrest = ifelse(Arrest == "TRUE",1,0),
    Domestic = ifelse(Domestic == "TRUE",1,0)    
  )

Lm_test_dataset <- test %>%
  group_by(District, Arrest,Domestic) %>%
  summarise(count = n()) %>%
  mutate(
    Arrest = ifelse(Arrest == "TRUE",1,0),
    Domestic = ifelse(Domestic == "TRUE",1,0) 
  )

```

### Linear Regression Summary of Train Dataset

```{r}
Model1 <- (lm(count ~ District * Arrest * Domestic , data = Lm_train_dataset))
summary(Model1)
```
From the above summary of the linear regression model for train dataset developed by us you can see that the R squared value is 0.84. that means our model is 84% fit for predicting the crime rate.
The is not much difference in the R-square value and the Adjusted R-squared Value.
Also the Pvalue is below the significant level (0.05)
This model explains a lot of variation within data and is significant. Thus this model is best fit for Predicting the Crime rate of Chicago.

### Linear Regression Summary of Test Dataset

```{r}
Model2 <- (lm(count ~ District * Arrest* Domestic  , data = Lm_test_dataset))
summary(Model2)
```
From the above summary of the linear regression model for test sample dataset developed by us you can see that the R squared value is 0.84. that means our model is 84% fit for predicting the crime rate.
The is not much difference in the R-square value and the Adjusted R-squared Value.
Also the P-value is below the significant level (0.05)
This model explains a lot of variation within data and is significant. Thus this model is best fit for Predicting the Crime rate of Chicago.

### Residual Plotting of the models

```{r,fig.show="hold", out.width="50%", results='hide', comment = " ", }
Model1_res <- resid(Model1)
Model2_res <- resid(Model2)

plot(fitted(Model1), Model1_res) + 
  abline(0,0)+
  title("PLOT - 01- Residual Of TRAIN dataset")

plot(fitted(Model2), Model2_res) + 
  abline(0,0)+
  title("PLOT - 02- Residual of TEST dataset")
```

<ul>
<li>The response variable for this model was Crime rate and the predictor variables were District ,Domestic and Arrests</li>
<li>observations are evenly distributed</li>
</ul>

### Q-Q PLOT

```{r,fig.show="hold", out.width="50%", results='hide', comment = " "}
qqPlot(Model1)+
  title("Q-Q PLOT MODEL 1")

qqPlot(Model2)+
  title("Q-Q PLOT MODEL 2")
```

From plot of model 1 and model 2 we can wee that most of the observation follow the normal line and there are less outliers




### MULTI COLLINEARITY
```{r, comment = " ", }
## model 1
vif(Model1)

## model 2
vif(Model2)
```
Multicollinearity help us to know if the predictor variables (independent) have correlation or not.
If the value is below 5 we can say that there is not correlation between the Predictor variables.









## Logistic Regression

We developed 2 model, one with variables having highest correlation and one with variables having lowest correlation to learn about difference between the best fit and worst fit model

```{r}

logistic_dataset_train <- train %>%
  group_by(Primary.Type,Domestic,Arrest,District) %>%
  summarise(count = n())%>%na.omit()

logistic_dataset_test <- train %>%
  group_by(Primary.Type,Domestic,Arrest,District) %>%
  summarise(count = n())%>%na.omit()

model3 <- glm(Arrest ~ .,data = logistic_dataset_train, family = "binomial")
summary(model3)

model4 <- glm(Arrest ~. ,data = logistic_dataset_test, family = "binomial")
summary(model4)

```

<ul>
<li>The intercept for our Model1 and model2 are 4.39 and </li>
<li>The Arrests decreases by 3.04 if the case type is Homicide as compared to Arrests made in Assault Cases</li>
<li>The Arrests decreases by 2.15 if the case type is Robbery</li>
<li>The Arrests decreases by 2.90 if the case type is Robbery</li>
<li>The Arrests decreases by 1.41 if the case is of Domestic violence</li>
</ul>

###The Ratio of computed odds ratio:

For showing the table with coefficients on the odds scale we have used the stargazer function. 

```{r, comment= " ", out.width="50%"}
coef_Model1 <-stargazer(model3,odd.ratio = TRUE ,type = "text")
coef_Model2 <-stargazer(model4,odd.ratio = TRUE ,type = "text")
```

The computed odds ratio of the Module generate using test and train data are similar.


### Effect Plots

The effect plots show us how predicted probabilities change when we change our independent variables.

```{r, fig.width= 20}

plot(allEffects(model3))

```

From the above plotting we can see the effects in arrest been made based on Type of Crime, Domestic violence, Districts wise and counts.
<ul>
<li>From 1st plot  you can see that the arrest been made in cases like Violent is more than rest of the cases.</li>
<li>From 2nd plot  you can see that the arrest been made in cases When domestic violence is presence is less.</li>
<li>From 3rd plot  you can see that the arrest been made in cases When the district cases(from 1 to 10) is in decreasing trend.</li>
<li>From 4th plot  you can see that the arrest been made in cases When the cases increases the arrests made are decreasing.</li>
</ul>


reference :
<ul>
<li> Multiple and Logistic Regression in R [DataCamp]</li>
<li>StatQuest: Logistic Regression. (2018, March 5). YouTube. https://www.youtube.com/watch?v=yIYKR4sgzI8&list=PLblh5JKOoLUKxzEP5HA2d-Li7IJkHfXSe</li>
<li>» Interpreting Residual Plots to Improve Your Regression. (2021, April 12). Qualtrics. https://www.qualtrics.com/support/stats-iq/analyses/regression-guides/interpreting-residual-plots-improve-regression/</li>
<li>Residual plots | Exploring bivariate numerical data | AP Statistics | Khan Academy. (2017, July 12). YouTube. https://www.youtube.com/watch?v=VamMrPZ-8fc</li>
<li>Z. (2021, April 22). How to Create a Residual Plot in R. Statology. https://www.statology.org/residual-plot-r/ </li>
<li>Visualizing two variables - Characterizing bivariate relationships. (n.d.). DataCamp. https://campus.datacamp.com/courses/correlation-and-regression-in-r/visualizing-two-variables?ex=4</li>
</ul>