---
title: "Group1"
author: "Rahul Jadhav"
date: "1/26/2022"
output: 
  html_document:
     theme: cerulean
     toc: true
     toc_float: true
---

```{r ,include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)

packages<-(c("dplyr","psych","ggplot2","vtable","maps","fastDummies","broom","gtools", "scales"))
package.check <- lapply(
  packages,
  function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
lapply(c("dplyr","psych","ggplot2","vtable","maps","fastDummies","broom"),require, character.only = TRUE)

```
## DATA PREPARATION

For cleaning the dataset, We have dropped all the rows with NA and there where columns with same name but different upper and lower alphabets, so we have corrected that.
We have even changed the class of Date variable from Character to Date so that we can later on use it for analysis.

```{r, comment = ""}
library(dplyr)
main_dataset <- read.csv('Chicago_Crimes_2001_to_2017_Part2.csv')
str(main_dataset)

## check for variables with NA's and drop NA's

sapply(main_dataset, function(x) sum(is.na(x)))

main_dataset <- na.omit(main_dataset)

## convertion of few variables and data variable class from character to POSIXct (data-time)
main_dataset <- main_dataset %>%
  mutate(
    Arrest = ifelse(
      Arrest == "False", "FALSE", 
      ifelse(
        Arrest == "True", "TRUE", main_dataset$Arrest
      )),
    Domestic = ifelse(
      Domestic == "False", "FALSE", 
      ifelse(
        Domestic == "True", "TRUE", main_dataset$Domestic
      )),
    Date = as.POSIXlt(Date, format = "%m/%d/%Y %I:%M:%S %p"),
    YearOfRegistered = format(main_dataset$Date, format="%Y"),
    Month = format(main_dataset$Date, format="%m"),
    Day = format(main_dataset$Date, format="%d"),
    Time = format(main_dataset$Date, format="%I")
    )

## subsetting categorical and numerical data to perform initial analysis
num_dataset <- main_dataset[sapply(main_dataset, is.integer) | sapply(main_dataset, is.numeric) ]
character_dataset <- main_dataset[sapply(main_dataset, is.character)]
character_dataset[sapply(character_dataset, is.character)] <- lapply(
  character_dataset[sapply(
    character_dataset, 
    is.character)]
  ,as.factor)
```

## SUMMARY STATISTIC

For numerical variable, we have created subset of numerical variable and categorical variable.


```{r}
psych::describe(num_dataset) %>%
  dplyr::select(count = n,
         Mean = mean,
         Standard_Deviation = sd,
         )%>%
  mutate(
    Coefficient_Variation = Standard_Deviation/Mean
  )

summary(character_dataset)

```


## BASIC ANALYSIS
### Primary type of crimes 

```{r}
library(dplyr)
library(ggplot2)
group_by_primary_type <- main_dataset %>%
  group_by(Primary.Type) %>%
  summarise(count = n())%>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange(desc(percentage)) %>%
  head(10)


ggplot(group_by_primary_type,aes(x=Primary.Type,y=count))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(label = comma)+
  ggtitle("Bar Plot : 01 Top 10 crimes reported in Chicago")

group_by_primary_type <- main_dataset %>%
  group_by(Primary.Type,Arrest) %>%
  summarise(count = n())%>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange(desc(percentage)) 


ggplot(group_by_primary_type,aes(x=Primary.Type,y=count,fill= Arrest))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(label = comma)+
  ggtitle("Bar Plot : 02 Crimes reportd in chicago with Arrested Data")


```

From the above plot you can see that from the from the amount of cases registered, the arrests are less.
Among all the aother primary type of crimes , Theft cases are reported at the highest followed by battery and criminal damage.

### Year-Wise Crime rate
```{r}
library(dplyr)
library(ggplot2)
group_by_Year <- main_dataset %>%
  group_by(Year,Arrest) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  arrange((Year))


ggplot(group_by_Year,aes(x=Year,y=count, fill= Arrest))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust=1))+
  ggtitle("Bar Plot : 03 Yearly Crime rate of Chicago from 2001 to 2017")



```
From the above plot you can see that between the year 2006 to 2010, the crime reported where highest.
After 2010 we can see the down trend in the crime rates

### ARREST
```{r}
group_by_Arrest <- main_dataset %>%
  group_by(Arrest)  %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  ) %>%
  mutate(ypos = cumsum(percentage)+ 0.2*percentage )
ggplot(group_by_Arrest, aes(x="", y=percentage, fill=Arrest)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme(legend.position="none") +
  theme_void()+
  ggtitle("Pie chart of Arrest")+ 
  geom_text(y = group_by_Arrest$ypos,label=group_by_Arrest$percentage)+
  ggtitle("Pie Chart : 01 Arrests among the total Crime reported")
```
From the above plot you can see that about 71.74% of the crime reported had no arrests where as 28.26% of the crime reported have arrests.


```{r}
group_by_District <- main_dataset %>%
  group_by(District,Arrest) %>%
  summarise(count = n()) %>%
  mutate(
    porportion = round(count/sum(count),2),
    percentage = round((count/sum(count)*100),2)
  )
ggplot(group_by_District,aes(x=District,y=count, fill= Arrest))+
  geom_bar(stat="identity")+
  ggtitle("Bar Plot : 05 District wise cases")+
  xlab("District")

```




## CORRELATION
Checking correlation between District crime primary type and arrests

```{r, fig.width=20, fig.height=15}

## from our eda we will concentrate on 9 primary type criminal cases and merge remaining to other primary type
primary_Type <- main_dataset %>%
  mutate(
    Primary.Type = ifelse(
      Primary.Type == "THEFT" , "THEFT", 
      ifelse(
        Primary.Type == "BATTERY", "BATTERY",
        ifelse(
          Primary.Type == "CRIMINAL DAMAGE", "CRIMINAL DAMAGE	",
          ifelse(
            Primary.Type == "NARCOTICS", "NARCOTICS",
            ifelse(
              Primary.Type == "ASSAULT", "ASSAULT",
              ifelse(
                Primary.Type == "BURGLARY", "BURGLARY",
                ifelse(
                  Primary.Type == "MOTOR VEHICLE THEFT", "MOTOR VEHICLE THEFT",
                  ifelse(
                    Primary.Type == "ROBBERY", "ROBBERY",
                    ifelse(
                      Primary.Type == "DECEPTIVE PRACTICE", "DECEPTIVE PRACTICE", "OTHER OFFENCES"
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  )

primary_Type <- primary_Type %>%
  select(Primary.Type,District,Arrest)
primary_type <- dummy_cols(primary_Type)


primary_type <- primary_type[-c(1,3)]
names(primary_type) <- c('District', 'ASSAULT', 'BATTERY', 'BURGLARY', 'CRIMINAL DAMAGE', 'DECEPTIVE PRACTICE','MOTOR VEHICLE THEFT', 'NARCOTICS', 'OTHER OFFENCES', 'ROBBERY', 'THEFT', 'NOT.Arrested', 'Arrested')
class(primary_type$Primary.Type)
ggcorrplot::ggcorrplot(cor(primary_type),
                       lab_size = 10,
                       tl.cex = 20,
                       title = "Correlation Plot:01 Correlation table"
                       ,type = "lower",
                       lab = TRUE)

```


## CHI-SQUARE ANALYSIS
Performing Chi-Square test to examine if Sale Price varies by Neighborhood

H0 (null hypothesis)        -> There is no relationship between District and Primary type of Crime
H1 (Alternative Hypothesis) -> There is relationship between District and Primary type of Crime
```{r}
chi <- main_dataset
myChisq <- xtabs(~ as.factor(Primary.Type) + as.factor(District), data = chi)

tidy(chisq.test(myChisq))%>%
  select(
    method,
    DF = parameter,
    p.value,
    statistic
  )%>%
  mutate(
    statistic = round(statistic,2)
  )
```

Interpretation :
A chi-square test was performed to examine the relation between District and Primary type of crimes in Chicago. The relation between these variables was significant, X^2 (1,N=792) = 297892.5, p = 0. 

Since the p.value is very significant, we will reject the null hypothesis.


## 2 WAY ANOVA TEST 
Performing ANOVA test to examine if District varies by Primary type of crime and Arrests.

H0(null hypothesis) -> 
  * Primary type of crime and Arrest has no effect on District. Also there is no interaction affect on District due to Primary type of crime and Arrests.
  
H1(Alternative Hypothesis) ->
  * Primary type of crime and Arrest has effect on District. Also there is interaction affect on District due to Primary type of crime and Arrests.

```{r }
two_way_fit <- aov(District ~as.factor(Primary.Type) * as.factor(Arrest) , data=main_dataset)
tidy(two_way_fit)%>%
  mutate(
    sumsq = signif(sumsq, digits = 2),
    p.value = round(p.value,3),
    statistic = round(statistic,3)
    
  )

```

Interpretation :
A two-way ANOVA was performed to analyze the effect of Primary type of crime and Arrest on District.
A two-way ANOVA revealed that -
Simple main effects analysis showed that Primary type of crime did have a statistically significant effect on District (p = 0).

 
Simple main effects analysis showed that Arrest did have a statistically significant effect on Sale Price (p = 0).

## INTERACTION PLOT FROM ANOVA


```{r, comment = " ", fig.width=22, fig.height=7, fig.align='center'}
## merging Exterior values lesser than 60 
test <- main_dataset
interaction.plot(x.factor = main_dataset$Primary.Type,
                 trace.factor = main_dataset$Arrest,
                 response = main_dataset$District,
                 fun = mean,
                 type = "b",
                 ylab = "District",
                 xlab = "Primary type of cases",
                 col = c("red","green"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "Arrest",
                 pch=c(19),             ### Symbols for levels of trace var.
                 fixed=TRUE,                    ### Order by factor order in data
                 leg.bty = "o")
 
```

## CONCLUSION

<> Based on initial analysis, the crime of Theft , Battery and criminal violence in more in Chicago

<> There is a decline trend in crimes after the year 2010.

<> From the cases reported only in 28.26% of the cases arrests are been carried out.

<> From chi square test we can see that there is a dependency of crime with district.

<> From anova test we can say that Primary type of crime and Arrest has effect on District. Also there is interaction affect on District due to Primary type of crime and Arrests
