---
title: "Module_4_Logistic"
author: "Rahul Jadhav"
date: "2/8/2022"
output:
  html_document:
     theme: cerulean
     toc: true
     toc_float: true
---

```{r setup, include=FALSE, comment = " "}
knitr::opts_chunk$set(echo = TRUE)
## This code checks if the package is installed or not, if not installed it will install them
packages<-(c("TOC","dplyr","gtools","broom","knitr","car","pacman","stargazer","effects","ggplot2"))
package.check <- lapply(
  packages,
  function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
lapply(packages,require, character.only = TRUE)

```


```{r}
main_dataset <- read.csv("AmesHousing.csv")

##Creating subset of relevant continuous variable
cleaned_dataset <- main_dataset[sapply(main_dataset, is.integer) | sapply(main_dataset, is.numeric) ]
cleaned_dataset <- cleaned_dataset %>%
  dplyr::select(-c(ï..Order,
                    PID,
                    Year.Remod.Add,
                    Garage.Yr.Blt,
                    Mo.Sold,
                    Yr.Sold
                    ))

```

## INTRODUCTION
In this assignment we are going to perform logistic Regression in Ames Dataset. In Machine Learning, Logistic Regression is commonly used to address categorization issues. When the answer response is a binary(1 and 0 )or 2 level factor variable such as yes or no, Male or Female, etc. and can be converted into binary, Logistic Regression is used.

The way Logistic regression differs from linear regression is, in linear regression you can use continuous variable and in Logistic regression you use Binary variable as your response variable. Also, A straight line is the best fit line in linear regression, and an S-shaped line, sometimes known as a "Squiggle," is the best fit line in logistic regression.

The coefficient of logistic regression can interpreted using 3 scales, Probability scale, Odds Scale and Log Odds Scale.

## SUMMARY TABLE

```{r}
psych::describe(cleaned_dataset) %>%
  dplyr::select(count = n,
         Mean = mean,
         Median = median,
         Standard_Deviation = sd,
         )%>%
  mutate(
    Coefficient_Variation = Standard_Deviation/Mean
  )
```

## PREPARING DATASET

For logistic regression we will convert selling price of the house into 1 and 0. For making it more sensible we will create a new variable called binary_sp and if the selling price is above 200,000 as 1 and 0 if selling price is lesser than 200,000.
```{r}
sapply(cleaned_dataset, function(x) sum(is.na(x)))
cleaned_dataset$binary_sp <- ifelse(cleaned_dataset$SalePrice > 200000,1,0) 
table(cleaned_dataset$binary_sp)
```

## DEVELOPING MODEL AND ITS INTERPRETATION

```{r}

Model1 <- glm(binary_sp ~ Overall.Qual + 
                Gr.Liv.Area + 
                Year.Built + 
                Fireplaces , data= cleaned_dataset , family = "binomial")
summary(Model1)
exp(coef(Model1))
```

<ul>
<li>The intercept for our Model is 1.91</li>
<li>The sale price increases by 3.83 by every increase in rating Quality of the house</li>
<li>The Sale Price increases by 1.00 by every increase in the General living area sqaure feet</li>
<li>The sale Price increases by 1.03 by every increase in the build year of the house</li>
<li>The sale Price increases by 2.44 by every increase in the Pool Area of the house</li>
</ul>

###The Ratio of computed odds ration:

For showing the table with coefficients on the odds scale we have used the stargazer function. 

```{r, comment= " "}
coef_Model1 <-stargazer(Model1,odd.ratio = TRUE ,type = "text")

```

### Augmented Logistic Regression Model:


```{r, comment= " "}
augment_model1<- augment(Model1, type.predict = "response") %>%
  mutate(.fitted = round(.fitted, 3), .resid = round(.resid, 3),
         .hat=round(.hat, 3), .cooksd = round(.cooksd, 3),
         price_hat = round(.fitted))%>%
  dplyr::select(c("binary_sp", ".fitted", ".resid", "price_hat"))
augment_model1
```

Broom package provides us with a function called Augment. Augment function provides us with important information about our model like the Fitted Value , residual value and price_hat

## Visualization :

### SalePrice and Overall Quality of House:

```{r, comment= " "}
ggplot(cleaned_dataset, aes(x= Overall.Qual, y=binary_sp))+
  geom_jitter(width = 0, height = 0.05)+
  scale_x_continuous(breaks = 1:10)+
  geom_smooth(method = "glm", method.args = list(family="binomial"), color = "blue")+
  xlab("Over all Quality of House")+
  ylab("Binary Selling Price")

```

S-shaped line helps us to determine th changing trend of the probability of scale of 0 and 1 with 1 unit change in the predictor variables.
From the Above plot we can see that the Selling Price of the house is 200000 dollar if the quality of the house is above 7. 


### SalePrice and General Living Area

```{r, comment= " "}
ggplot(cleaned_dataset, aes(x= Gr.Liv.Area, y=binary_sp))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(method = "glm", method.args = list(family="binomial"), color = "blue")+
  xlab("General Living Area")+
  ylab("Binary Selling Price")

```


From the Above plot we can see that the Selling Price of the house is 200000 dollar if the General Living Area is Above 2000 sq feet.

### SalePrice and Year of built

```{r, comment= " "}
ggplot(cleaned_dataset, aes(x= Year.Built, y=binary_sp))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(method = "glm", method.args = list(family="binomial"), color = "blue")+
  xlab("Year")+
  ylab("Binary Selling Price")

```

From the Above plot we can see that the Selling Price of the house is 200000 dollar if the year of the house is more than 1980. 

### SalePrice and No of fire Places in the house

```{r, comment= " "}
ggplot(cleaned_dataset, aes(x= Fireplaces, y=binary_sp))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(method = "glm", method.args = list(family="binomial"), color = "blue")+
  xlab("No of FirePlaces")+
  ylab("Binary Selling Price")

```

From the Above plot we can see that the Selling Price of the house is 200000 dollar if the house has fire place. 

### Effect Plots

The effect plots show us how predicted probabilities change when we change our independent variables.

```{r, fig.width= 20}

Model2 <- glm(binary_sp ~ Overall.Qual + 
                Gr.Liv.Area + 
                Year.Built + 
                Fireplaces +
                Overall.Cond+
                Garage.Cars

                , data= cleaned_dataset , family = "binomial")
plot(allEffects(Model2))

```

## References:
<ul>
<li> Multiple and Logistic Regression in R [DataCamp]</li>
<li>StatQuest: Logistic Regression. (2018, March 5). YouTube. https://www.youtube.com/watch?v=yIYKR4sgzI8&list=PLblh5JKOoLUKxzEP5HA2d-Li7IJkHfXSe</li>
</ul>